---
title: "Mini-Project 2"
author: "Ester Zhao, Lizette Carpenter, and Emily Kim"
date: "October 31, 2017"
output: html_document
---

```{r}
library(tidyverse)
library(dplyr)

```

Instructions:
1. discuss an interesting and well-motivated topic
2. involve some non-trivial data wrangling (e.g. not just a bunch of mutate()s)
3. provide a well thought-out, informative analysis
4. convey some sort of insight
5. be well-written


## Loading the data

```{r, include=FALSE}
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
```

```{r}
cand_info <- candidates %>%
  select(cand_id, cand_name, cand_party_affiliation, cand_office_state, cand_election_yr) %>%
  filter(cand_office_state == "US") %>%
  arrange(cand_election_yr)


cand_contributions <- contributions %>%
  select(name, cand_id, transaction_amt) %>%
  filter(cand_id != "") %>%
  arrange(cand_id)

total_cand_cantrib <- cand_contributions 
  
  
```


Plans:
    1. Average total contributions by political party (presidential election)
        - bar graph (area for contribution)
    2. Contribution amount and election results (scatterplot/line graph)

Steps:
    1. Find a way to eliminate NA data from dataset DONE
    2. Create a new data set that contains total contributions for each candidate
        - Create a list of all cand_id
        - Create a function that runs through the list of names & checks in cand_contributions to consolidate cand total contributions
    3. Merge total_contributions with cand_info using a join() function
    4. Create a cool data graphic!!
    5. Get an A!!!
